|

### 2020-03-02 analysis/paper/manuscript.Rmd
Add a few extra lines to the conclusions and tweak the intro

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and test custom parameterizations of a Von Bertalanffy Growth model [@catalan2018; see Appendix A for more examples]. While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In both cases interactive visuals are nested inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)).

@@

We must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming, but all had their formal training in field-based population ecology. While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches. There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/); a benthic-focused tool to assist with marine license application and compliance monitoring.
- Species Dashboard (https://shiny.marine.ie/speciesdash/); a tool for exploring biological fisheries data collected in Irish waters
- The Stock Book (https://shiny.marine.ie/stockbook/); a tool for accessing up-to-date information on the status of fish stocks in Irish waters.
- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.
- EcoCast Explorer (https://coastwatch.pfeg.noaa.gov/ecocast/explorer.html); an application for displaying results from a dynamic ocean management tool focused on bycatch avoidance [see @hazen2018 for more detail].
- B-VBGM! Bayesian Von Bertalanffy Growth Model tool (https://fishecology.shinyapps.io/B-VBGM/); a supplement Catalán et al. [-@catalan2018] that facilitates the exploration of the model described in the paper.
- Beyond Temperature tool (https://heatherwelch.shinyapps.io/beyond_temperature/); a supplement to McHenry et al. [-@mchenry2019] that allows the exploration of habitat suitability hindcasts and projections by species and model type.
- BRSApp (https://catalinagomez.shinyapps.io/BRSApp/); a supplement to Gomez et al. [-@gomez2016] that allows the exploration of behavioural response severity scores of marine mammals to noise.
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice (developed to meet an "EU request on dissemination of ICES advice beyond pdf files").

>>>>

Fisheries scientists are a part of this trend and, as such, there are a growing number of interactive tools being released that facilitate the exploration and communication of fisheries science. For instance, they can be used to disseminate real-time information for dynamic ocean management [@hazen2018; @maxwell2015], explore hindcasts and projections from various habitat suitability models [@mchenry2019], and evaluate different fisheries management approaches [@miller2019; @goethel2019] (see Appendix A for more examples). While few of these tools completely replace static modes of presenting scientific information, they allow audiences to explore the data and models in ways not amenable to static formats and, in the process, they improve accessibility and augment communication efforts. These are important tools to take advantage of as we proceed through an age of heightened scientific scrutiny [@mcinerny2014], and here we aim to promote their use in fisheries science by demonstrating a series of interactive visualization tools built to aid the peer review of stock assessments.

We first demonstrate the application of interactive tools to data that are commonly presented in stock assessments to show how they can simplify the detailed exploration of data from long-term monitoring programs. Second, we focus on the modeling aspect of stock assessment and demonstrate how interactive tools can be used to explore, diagnose and communicate results from an integrated assessment model. This structure corresponds to two important  steps in the stock assessment process: 1) data exploration, and 2) data modeling. In both cases interactive visuals are nested inside dynamic documents called "dashboards". The concept of using a dashboard was borrowed from the business community where they are frequently used to group a series of interactive visuals and tables to provide at-a-glance views of key performance indicators. These dashboards were constructed within the framework of the R programming language [@R] and the RStudio IDE [@Rstudio] using three or more of the packages listed in Table (Table \@ref(tab:pkg)). See Supplement 1 for a "getting started" tutorial that highlights one approach to constructing a dashboard using R; using this as a guide, an R user can start constructing basic dashboards within a day.

@@

Visualizations play an important role in communicating fisheries science to a wide range of stakeholders [@levontin2017] and here we posit that interactive tools extend their utility by providing a richer approach to communicating scientific information. However, we must acknowledge there are costs to adopting the use of interactive tools. The first and obvious one being that staff are required to have both the time and training to effectively develop these tools. Ideally, professional programmers and graphic designers would be hired to construct dashboards and other interactive info-graphics [@mcinerny2014], but fisheries management organizations rarely have the resources to fund such initiatives. Consequently, the process of communicating fisheries science tends to be the responsibility of fisheries scientists who lack formal training in these fields. In our three examples, the developers (first, second and third authors) were relatively proficient with R and have natural aptitudes for programming and graphic design, but all had their formal training in field-based population ecology. While some learning time was needed (a few weeks), the learning curve was relatively minor because there was no need to learn a new programming language (e.g. JavaScript, CSS, etc.). Moreover, a growing user base means that most programming issues were readily addressed through simple internet searches. There are also a growing number of exemplars (Appendix A), beyond what we show here, to inspire the development of new interactive tools. We expect the accessibility and quality of these tools to continue to improve as more people in the fisheries community develop and use interactive documents. It is our experience that the upfront investment in time has already paid dividends in terms of efficiencies of delivering products for subsequent stock assessments, and in the quality of the data exploration and understanding.

@@

- Marine Aggregates Application (https://openscience.cefas.co.uk/ma_tool/); a benthic-focused tool to assist with marine license application and compliance monitoring.
- Species Dashboard (https://shiny.marine.ie/speciesdash/); a tool for exploring biological fisheries data collected in Irish waters
- The Stock Book (https://shiny.marine.ie/stockbook/); a tool for accessing up-to-date information on the status of fish stocks in Irish waters.
- Data-Limited Fisheries Toolkit demo (https://www.datalimitedtoolkit.org/demo); an application for demonstrating the key features of the management strategy evaluation toolkit.
- Toy Tuna MSE application (https://puntapps.shinyapps.io/tunamse/); a toy MSE example built to allow users to enter various management procedures and assess their performance.
- Exploring Groundfish in an Ecosystem Context (https://incorporatingecosystemapproach.shinyapps.io/indiapp/); a tool accessing results from an analysis exploring links between stock indices and ecosystem trends.
- EcoCast Explorer (https://coastwatch.pfeg.noaa.gov/ecocast/explorer.html); an application for displaying results from a dynamic ocean management tool focused on bycatch avoidance [see @hazen2018 for more detail].
- B-VBGM! Bayesian Von Bertalanffy Growth Model tool (https://fishecology.shinyapps.io/B-VBGM/); a supplement Catalán et al. [-@catalan2018] that facilitates the exploration of the model described in the paper.
- Beyond Temperature tool (https://heatherwelch.shinyapps.io/beyond_temperature/); a supplement to McHenry et al. [-@mchenry2019] that allows the exploration of habitat suitability hindcasts and projections by species and model type.
- BRSApp (https://catalinagomez.shinyapps.io/BRSApp/); a supplement to Gomez et al. [-@gomez2016] that allows the exploration of behavioural response severity scores of marine mammals to noise.
- VISA tool (e.g. https://ices-tools-dev.github.io/VISA_tool/hke.27.3a46-8abd.html#ices_advice_2018); an alternative to pdf files for displaying and exploring up-to-date ICES advice (developed to meet an "EU request on dissemination of ICES advice beyond pdf files").
